Long name of the paper

### 2018-10-05 anzjs2018/sections/03_models.tex
network model

% The second part of this appraoch involves a network model of transit network travel times along links in the network.
% Reliable ETAs will require a historical data component,
% used as a prior in the absense of data as well as for
% predicting future states, 
% and a realtime update component where the travel times of vehicles
% are used to update the current estimate of network state.

% The model will be another recursive Bayesian model as used for the vehicle state,
% as this provides a simple way to model realtime data.
% Since travel times can safely be assumed to be unimodal,
% we use an information filter (IF),
% which is a transformation of the Kalman filter
% using information instead of covariance,
% and allows the simple combination of multiple observations,
% for example when multiple buses traverse a road in the same update.
% The prediction step will involve a transition function that incorporates historical data
% and the update step will weight between observed and predicted travel times.

% The network state $\Theta_c = \{\theta_c^j\}_{j = 1}^J$ is the travel time 
% of transit vehicles along road segment $j$ at time $t_c$.
% In this implementation, we are assuming independence between each segment
% to provide a simple basis for the model; of course in reality this
% is not true and future work will investigate ways of incorporating this.

% Using an IF requires a transition matrix, $\mathbf{F}_c$;
% however, in this case the travel time model is constant so no
% transition matrix is required (i.e., $\mathbf{F}_c = 1$).
% Let $\Delta_c = t_c - t_{c-1}$ and $P_c^j$ the state uncertainty
% with system noise $Q_c = \Delta_c \sigma_b^2$, then the update equations are
% simply the normal KF update equations,
% \begin{align}
% \label{eq:kf_transition}
% \hat\theta^j_{c|c-1} &= \hat\theta^j_{c|c-1} \\
% P^j_{c|c-1} &= P^j_{c-1|c-1} + Q_c
% \end{align}

% For the update step, however, we need to transform into an information
% space parameterised by the information matrix 
% $\mathbf{Z}^j_{c|c-1} = P_{c|c-1}^{-1}$ 
% and the information vector $\mathbf{\hat z}^j_{c|c-1} = \hat\theta_{c|c-1} P_{c|c-1}^{-1}$.
% The measurement data obtained from the vehicle model are
% the travel time of vehicle $m$ along segment $j$,
% $\bar b^m_j$, and the uncertainty $s^m_j$.
% These can be transformed to a measurement information covaraince matrix
% and vector using the measurement matrix $\mathbf{H} = 1$ since,
% in this case, the observed state is the underlying state we are estimating (travel time).
% \begin{equation}
% I^m_{jc} = \frac{1}{(s^m_j)^{2}}\quad\text{and}\quad
% i^m_{jc} = \frac{\bar b^m_j}{(s^m_j)^2}
% \end{equation}
% The information update is now the sum of the information for all vehicles
% that traversed the segment since the last update, so
% \begin{align*}
% \mathbf{Z}^j_{c|c} &= \mathbf{Z}^j_{c|c-1} + \sum_{m=1}^M I^m_{jc} \\
% \mathbf{\hat i}^j_{c|c} &= \mathbf{\hat z}^j_{c|c-1} + \sum_{m=1}^M i^m_{jc}
% \end{align*}
% which are easily transformed back to get the desired state parameters.
% \begin{equation}
% \hat\theta^j_{c|c} = \frac{\mathbf{\hat z}^j_{c|c}}{\mathbf{Z}^j_{c|c}} 
% \quad\text{and}\quad
% P^j_{c|c} = \frac{1}{\mathbf{Z}^j_{c|c}}
% \end{equation}

>>>>

The second part of this appraoch involves a network model of transit network travel times along links in the network.
Reliable ETAs will require a historical data component,
used as a prior in the absense of data as well as for
predicting future states, 
and a realtime update component where the travel times of vehicles
are used to update the current estimate of network state.

The model will be another recursive Bayesian model as used for the vehicle state,
as this provides a simple way to model realtime data.
Since travel times can safely be assumed to be unimodal,
we use an information filter (IF),
which is a transformation of the Kalman filter
using information instead of covariance,
and allows the simple combination of multiple observations,
for example when multiple buses traverse a road in the same update.
The prediction step will involve a transition function that incorporates historical data
and the update step will weight between observed and predicted travel times.

The network state $\boldsymbol\theta_c = \{\theta_c^\ell\}_{\ell = 1}^L$ is the travel time 
of transit vehicles along road segment $\ell$ at time $t_c$.
In this implementation, we are assuming independence between each segment
to provide a simple basis for the model; of course in reality this
is not true and future work will investigate ways of incorporating this.

Since the road state model is very simple, 
no transition matrix or measurement matrix is required,
and the model from (\ref{eq:rbe_model}) reduces to
\begin{equation}
\begin{split}
\hat \theta_c^\ell &= \theta_{c-1}^\ell + \Delta_c^\ell v_c^\ell \\
Z_c^\ell &= \theta_c^\ell + e_c^\ell
\end{split}
\end{equation}
where $\Delta_c = t_c - t_{c-1}$,
$v_c^\ell \sim \mathrm{N}(0, \nu^2)$ is the system noise,
and $e_c^\ell \sim \mathrm{N}(0, R_c^\ell)$ is the measurement error
as estimated from the vehicle model using particle uncertainty.

The state is estimated by its mean $\hat \theta_c^\ell = \mathrm{E}(\theta_c^\ell)$
and uncertainty $P_c^\ell = \mathrm{var}(\theta_c^\ell)$,
which are predicted from the previous state estimate using the prediction
equations
\begin{align*}
\label{eq:kf_transition}
\hat \theta^\ell_{c|c-1} &= \hat \theta^\ell_{c|c-1} \\
P^\ell_{c|c-1} &= P^\ell_{c-1|c-1} + \Delta_c^\ell \nu_c^\ell
\end{align*}

For the update step, however, we need to transform into an information
space parameterised by the information matrix $U^\ell_c$
and the information vector $u^\ell_c$,
\begin{equation}
\label{eq:information_transform}
U^\ell_{c|c-1} = \frac{1}{P_{c|c-1}}
\quad\text{and}\quad
\hat u^\ell_{c|c-1} = \frac{\hat \theta^\ell_{c|c-1}}{P^\ell_{c|c-1}}
\end{equation}
The measurement data obtained from the vehicle model are
the travel time of vehicle $m$ along segment $\ell$,
$\bar z_c^{m\ell}$, and the uncertainty $s^{m\ell}_c$.
These can be transformed to a measurement information covaraince matrix
and vector using the measurement matrix,
\begin{equation*}
B^{m\ell}_c = \frac{1}{(s^{m\ell}_c)^{2}}\quad\text{and}\quad
b^{m\ell}_c = \frac{\bar z^{m\ell}_c}{(s^{m\ell}_c)^2}
\end{equation*}
The information update is now the sum of the information for all vehicles
that traversed the segment since the last update, so
\begin{align*}
U^\ell_{c|c} &= U^\ell_{c|c-1} + \sum_{m=1}^M B^{m\ell}_{c} \\
\hat u^\ell_{c|c} &= \hat u^\ell_{c|c-1} + \sum_{m=1}^M b^{m\ell}_{c}
\end{align*}
which are back transformed to get the updated state parameters
using the inverse of (\ref{eq:information_transform})
\begin{equation*}
\hat \theta^\ell_{c|c} = \frac{\mathbf{\hat u}^\ell_{c|c}}{\mathbf{U}^\ell_{c|c}} 
\quad\text{and}\quad
P^\ell_{c|c} = \frac{1}{\mathbf{U}^\ell_{c|c}}.
\end{equation*}
